---
title: "Covid-19 Positive Tests by Local Authority"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    runtime: shiny
---

<style>                     
.navbar {
  background-color:#00ad93;
  border-color:black;
}
</style> 

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tidyquant)
library(ggthemes)
library(ggdark)
library(DT)
library(plotly)
library(shiny)
library(httr)

today_date <- Sys.Date()

#tmp <- tempfile()
#GET("https://coronavirus.data.gov.uk/downloads/csv/coronavirus-cases_latest.csv", write_disk(tmp))
#readLines(tmp)
#df <- read.csv(tmp)
df <- read.csv('coronavirus-cases_latest.csv')

df <- df %>%
  mutate_at('Specimen.date', ~ as.Date(.)) %>%
  filter(Area.type == "Upper tier local authority") %>%
  select(Area.name, Area.code, Specimen.date, Daily.lab.confirmed.cases,
    Cumulative.lab.confirmed.cases, Cumulative.lab.confirmed.cases.rate) %>% 
  rename(
    area = 'Area.name',
    code = 'Area.code',
    date = 'Specimen.date', 
    daily_cases = 'Daily.lab.confirmed.cases', 
    cumulative_cases = 'Cumulative.lab.confirmed.cases', 
    cases_rate = 'Cumulative.lab.confirmed.cases.rate'
  ) %>%
  group_by(area) %>%
  tq_mutate(
    # tq_mutate args
    select = daily_cases,
    mutate_fun = rollapply, 
    # rollapply args
    width = 7,
    align = "right",
    FUN = mean,
    # mean args
    na.rm = TRUE,
    # tq_mutate args
    col_rename = "daily_cases_rolling_mean_7_days_per_la"
  ) %>%
 ungroup

df <- df %>%
  group_by(date) %>%
  summarise(daily_cases_rolling_mean_7_days_per_la_mean = mean(daily_cases_rolling_mean_7_days_per_la)) %>%
  ungroup %>%
  tq_mutate(
    # tq_mutate args
    select = daily_cases_rolling_mean_7_days_per_la_mean,
    mutate_fun = rollapply, 
    # rollapply args
    width = 7,
    align = "right",
    FUN = mean,
    # mean args
    na.rm = TRUE,
    # tq_mutate args
    col_rename = "daily_cases_rolling_mean_7_days_all_las"
  ) %>%
  select(-daily_cases_rolling_mean_7_days_per_la_mean) %>%
  right_join(df) %>%
  mutate_at(
    c(
      "daily_cases_rolling_mean_7_days_per_la", 
      "daily_cases_rolling_mean_7_days_all_las"
    ), 
    round)

# I need this so that filtering doesn't return zero-row data frames
# Ideally, data will be automaticall y download, so this piece of code is temp only
if (today_date > max(df$date)) {
  today_date <- max(df$date)
}

cases_change <- df %>%
  filter(
    date >= today_date - 9, 
    date <= today_date - 3
  ) %>%
  group_by(area) %>% 
  summarise(cases_recent_week_within = 
    sum(daily_cases, na.rm = TRUE)) %>%
  ungroup %>%
  full_join(
    df %>%
      filter(
        date >= today_date - 16, 
        date <= today_date - 10
      ) %>%
      group_by(area) %>% 
      summarise(cases_previous_week_within = 
        sum(daily_cases, na.rm = TRUE)) %>%
      ungroup
  ) %>%
  full_join(
    df %>% 
      filter(date <= today_date - 9) %>%
      group_by(area) %>% 
      summarise(cases_recent_week_cumulative = 
        sum(daily_cases, na.rm = TRUE)) %>%
      ungroup
  ) %>%
  full_join(
    df %>% 
      filter(date <= today_date - 16) %>%
      group_by(area) %>% 
      summarise(cases_previous_week_cumulative = 
        sum(daily_cases, na.rm = TRUE)) %>%
      ungroup
  ) %>%
  mutate(
    change_within = cases_recent_week_within - cases_previous_week_within,
    change_cumulative = 
      cases_recent_week_cumulative - cases_previous_week_cumulative
  )

make_hist <- function(df) {
  df %>%
    ggplot(
      aes(
        date, 
        daily_cases, 
        text = 
          paste(
            paste0('Local Authority: ', area),
            '\n',
            paste0('Day: ', date),
            '\n',
            paste0('Number of cases on that day: ', daily_cases)
          ),
        alpha = 0.5,
        group = area
      )
    ) +
    geom_bar(stat = "identity", position = "identity", fill = "#00ad93") +
    geom_line(aes(date, daily_cases_rolling_mean_7_days_per_la)) +
    geom_line(aes(date, daily_cases_rolling_mean_7_days_all_las), linetype = "dotted") +
    scale_colour_manual(
      values = c("daily_cases_rolling_mean_7_days_per_la" = "#39809E",
               "daily_cases_rolling_mean_7_days_all_las" = "red"), 
    name = NULL, 
    labels = "14-day rolling average"
   ) +
    theme_bw() +
    theme(
      legend.position = 'none',
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) + 
    xlab('') + 
    ylab('Daily cases')
}

make_bar <- function(df, key_var, chart_title, top_how_many = 20) {
  df %>%
    select(area, all_of(key_var)) %>%
    group_by(area) %>%
    summarize_all(max) %>%
    ungroup %>%
    rename(key_var = all_of(key_var)) %>%
    arrange(key_var) %>%
    mutate(area = factor(area, levels = area)) %>%
    arrange(desc(all_of(key_var))) %>%
    slice(1:top_how_many) %>%
    ggplot(
      aes(
        area, 
        key_var, 
        text = 
          paste(
            paste0('Local Authority: ', area),
            '\n',
            paste0('Number of cases: ', key_var)
          ),
        alpha = 0.5
      )
    ) +
    geom_bar(stat = "identity", width = 0.8, fill = '#00ad93') +
    coord_flip() +
    labs(x = "", y = "", title = chart_title) +
    theme_bw() +
    theme(
      legend.position = 'none',
      axis.title = element_text(size = 14, colour = "black"),
      axis.text.y = element_text(size = 11, face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
}
  
bar1 <- make_bar(
  df, 
  "cumulative_cases", 
  "Number of cases across the pandemic", 
  top_how_many = 20
)

recent_week_df <- df %>% 
  filter(date >= today_date - 9 & date <= today_date - 3) %>%
  group_by(area) %>% 
  summarise(cases = sum(daily_cases)) %>% 
  arrange(desc(cases))

chart_max <- round(max(recent_week_df$cases), -2)

bar2 <- recent_week_df %>%
  make_bar(
    "cases",
    paste(
      "Cases between", 
      format(today_date - 9,"%a %b %d"),
      "and",
      format(today_date - 3,"%a %b %d")
    )
  )

scatter_cumulative <- cases_change %>%
  ggplot(
    aes(
      x = cases_recent_week_cumulative, 
      y = change_cumulative,
      text = 
        paste(
          paste0('Local Authority: ', area),
          '\n',
         paste0('Number of new cases within time period: ', change_cumulative)
        )
    )
  ) +
  geom_point(colour = "#00ad93", fill = "#00ad93", 
    size = abs(cases_change$change_cumulative) / 50) +
  labs(
    title = 
      paste0(
        "Increase in cumulative cases between ", 
        format(today_date - 9, "%a %b %d"), 
        " and ", 
        format(today_date - 16, "%a %b %d")
     ),
    subtitle = "",
    caption = "",
    x = 
      paste0(
        "Cumulative cases up to ", 
        format(today_date - 9, "%a %b %d")
    ),
    y =  "Change in cases"
  ) +
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )

scatter_within <- cases_change %>%
  ggplot(
    aes(
      x = cases_recent_week_within, 
      y = change_within,
      text = 
        paste(
          paste0('Local Authority: ', area),
          '\n',
          paste0('Change in cases within time period: ', change_within)
        )
    )
  ) +
  geom_point(colour = "#00ad93", fill = "#00ad93", 
    size = abs(cases_change$change_within) / 50) +
  labs(
    title = 
      paste0(
        "Change in cases between week ", 
        format(today_date - 15, "%a %b %d"), 
        " - ",
        format(today_date - 9, "%a %b %d"), 
        " and week ", 
        format(today_date - 22, "%a %b %d"),
        " - ",
        format(today_date - 16, "%a %b %d")
      ),
    subtitle = "",
    caption = "",
    x = 
      paste0(
        "Cases in week ", 
        format(today_date - 15, "%a %b %d"),
        " - ",
        format(today_date - 9, "%a %b %d")
      ),
    y =  "Change in cases"
  ) +
  theme_bw() + 
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )
```

Daily Cases Over Time
=======================================================================
  
Column {.sidebar}
-----------------------------------------------------------------------
  
```{r}
selectInput("area", "Local Authority:", 
  choices = sort(unique(df$area)))
```

Column
-----------------------------------------------------------------------
  
### Daily positive Covid-19 tests by Local Authority
  
```{r}
renderPlot({
  df %>%
    filter(
      area == input$area
    ) %>%
    make_hist
})
```

Most Positive Tests 
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------


### Across Pandemic
```{r}
ggplotly(bar1, tooltip = c('text'))
```

### Last week 
```{r}
ggplotly(bar2, tooltip = c('text'))
```

Recent Change
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------


### Change in cumulative cases
```{r}
ggplotly(scatter_cumulative, tooltip = c('text'))
```

### Change in weekly cases
```{r}
ggplotly(scatter_within, tooltip = c('text'))
```

About
=======================================================================

**PHE Postive Tests by Local Authority**

Last updated: `r today_date`

To follow

**Data**
PHE coronvirus dashboard

**Using the plots interactively**

- Histogram of daily positive Covid-19 tests by Local Authority (LA)
  - At first view, the plot displays the histograms from all LAs on top of each other
  - For LA-specific histograms, double-click on the LA of choice on the legend on the right-hand side of the plot
  - Hovering over the histogram bars activates a tooltip with additional information
  - Hovering over the plot area activates a toolbar with additional options. The toolbar is a default feature of the software used to create the plot, so it may or may not prove useful

- Bar plots of total number of cases (whole period or last week) per LA
  - The 20 LAs with the highest number of cases are plotted
  - The swap between blue and orange bars is for making it easier to distinguish between consecutive bars
  - Hovering over the bars activates a tooltip with the actual number of cases in each LA
  - Hovering over the plot area activates a toolbar with additional options. The toolbar is a default feature of the software used to create the plot, so it may or may not prove useful
